---
title: "Survey Monkey - Data Transformation"
author: "Vish"
date: "24/08/2021"
output: html_document
---

```{r}
library(dplyr)#data manipulating
library(tidyr)#to gather data (melt)
library(readxl)#to read xlsx files
library(openxlsx)#write to excel 
```


```{r}
pwd <- getwd()
pwd
```
```{r}
data_import.T <- read_excel("Data - Survey Monkey Output Edited.xlsx" , sheet = "Edited_Data")
```

```{r}
dataset_modified.T <- data_import.T # Make a copy of the dataframe
colnames(dataset_modified.T)
```

```{r}
columns_to_drop.T <- c('Start Date', 'End Date', 'Email Address', 'First Name', 'Last Name', 'Custom Data 1')
columns_to_drop.T
```

```{r}
dataset_modified.T <- select(dataset_modified.T,-(columns_to_drop.T)) 
dim(dataset_modified.T)
```

```{r}
id_vars.T <- colnames(dataset_modified.T)[1:8] 
id_vars.T
```
Melts all column after the first 8 columns. Using ```gather``` from tidyr

```{r}
dataset_melted.T <- dataset_modified.T %>% 
  gather("Question.+.Subquestion", "Answer", -id_vars.T)


dim(dataset_melted.T)
```

```{r}
questions_import.T <- read_excel("Data - Survey Monkey Output Edited.xlsx", sheet="Question")
```

```{r}
questions.T <- questions_import.T
dim(questions.T)
str(questions.T)
```
Join two datasets tables using ```left_join``` from dplyr library.

```{r}
dataset_merged.T <- left_join(dataset_melted.T,questions.T, 
                              by= c("Question.+.Subquestion" = "Question + Subquestion"))
  
dim(dataset_merged.T)
```
Filter all na Answers. 

```{r}
respondents.T <- dataset_merged.T %>% 
  filter(!is.na(Answer))

dim(respondents.T)
```
To find the unique respondents for each question we use ```group_by``` followed by ```n_distinct``` on the Respondent ID

```{r}
respondents.T1 <- respondents.T %>% 
    group_by(Question) %>% 
    summarise(number_of_distinct_answers = n_distinct(`Respondent ID`))
    
dim(respondents.T1)
str(respondents.T1)
```
We now merge the two tables so the number of unique respondents are shown along side the question. 

```{r}
dataset_merged_two.T <- left_join(dataset_merged.T, respondents.T1, 
                                  by =c("Question"= "Question"))
  
dim(dataset_merged_two.T)
```

```{r}
same_answer.T <- dataset_merged.T %>% 
  filter(!is.na(Answer))
dim(same_answer.T)
```
To find the same answers for each question we use ```group_by``` on both Question+Subquestion and  Answer followed by ```n_distinct``` on the Respondent ID

```{r}
same_answer.T1 <- same_answer.T %>% 
 group_by(`Question.+.Subquestion`, Answer) %>% 
  summarise(number_of_same_answer = n_distinct(`Respondent ID`))
    

dim(same_answer.T1)
```
Now merge the same answer table with the dataset. We use the columns 'Question.+.Subquestion' and 'Answer'  as matching columns.

```{r}
dataset_merged_three.T <- left_join(dataset_merged_two.T, same_answer.T1, 
                                  by=c('Question.+.Subquestion', 'Answer'))
  
dim(dataset_merged_three.T)
```

```{r}
colnames(dataset_merged_three.T)
```
Renaming columns. 

```{r}
dataset_merged_three.T <- dataset_merged_three.T %>% 
  rename("Division" =`Identify which division you work in. - Response` ,
          "Division Other"=`Identify which division you work in. - Other (please specify)`,
          "Position"=`Which of the following best describes your position level? - Response` ,
          "Generation"=`Which generation are you apart of? - Response`,
          "Gender"=`Please select the gender in which you identify. - Response`,
          "Tenure"=`Which duration range best aligns with your tenure at your company? - Response`,
          "EmploymentType"=`Which of the following best describes your employment type? - Response`,
          "Respondents"=number_of_distinct_answers ,
          "SameAnswer"=number_of_same_answer)
```

Exporting transformed data. 

```{r}

write.xlsx(dataset_merged_three.T, paste(pwd , "/Final_Output_R_T.xlsx", sep=""))
```


